---
title: 'IO III: Entry Models'
output:
  pdf_document: default
  html_notebook: default
---

# 0. Libraries and set-up

First, the libraries used:

```{r}
# install.packages("haven")
library("haven") # to charge dta files
# install.packages("VGAM")
library("VGAM") # to get the bivariate normal
# install.packages("mlogit")
library(mlogit)
```

Now we remove everything from the environment and charge the dataset.

```{r}
rm(list = ls())
data = read_dta("jiadata.dta")
head(data)
```

# 3. Implementation

## 3.1. Two Types and Sequential Entry (Berry Model)

Mostly, translating Stata code to R.

Create market configurations variable and a constant:

```{r}
data["y"] = numeric(dim(data)[1]) + as.numeric((data$walmart==1)&(data$kmart==0)) + 2*as.numeric((data$walmart==0)&(data$kmart==1)) + 3*as.numeric((data$walmart==1)&(data$kmart==1))
data["comp"] = numeric(dim(data)[1]) + 1
```


### 3.1.1. Walmart moves first, binormal


### rho = 0.5, walmartbi_ext_pos

```{r}
walmartbi_ext_pos = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1]*data$population + par[2]*data$spc + par[3]*data$urban
  zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  zk <- par[7] + par[8]*data$midwest
  com <- par[9]*data$comp
  
  pik <- xb + zk
  piw <- xb + zw
  piduow <- xb + zw + com
  piduok <- xb + zk + com
  athrho = 0.5
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_duo - prob_w - prob_nofirm
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$y == 0, log(prob_nofirm), ifelse(data$y == 1, log(prob_w), ifelse(data$y == 2, log(prob_k), log(prob_duo))))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```


```{r}
optim(par=c(1.5, 1.5, 1.2, -11.8, -1.1, 0.72, -19.6, 0.37, -0.20), walmartbi_ext_pos, data_in=data, control = list("maxit"=10000), method="Nelder-Mead")
```


### rho = -0.5, walmartbi_ext_neg

```{r}
walmartbi_ext_neg = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1]*data$population + par[2]*data$spc + par[3]*data$urban
  zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  zk <- par[7] + par[8]*data$midwest
  com <- par[9]*data$comp
  
  pik <- xb + zk
  piw <- xb + zw
  piduow <- xb + zw + com
  piduok <- xb + zk + com
  athrho = -0.5
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_duo - prob_w - prob_nofirm
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$y == 0, log(prob_nofirm), ifelse(data$y == 1, log(prob_w), ifelse(data$y == 2, log(prob_k), log(prob_duo))))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```

```{r}
optim(par=c(1.5, 1.5, 1.2, -11.8, -1.1, 0.72, -19.6, 0.37, -0.20), walmartbi_ext_pos, data_in=data, control = list("maxit"=10000), method="Nelder-Mead")
```


## Estimating Rho, walmartbi_ext_est

```{r}
walmartbi_ext_est = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1]*data$population + par[2]*data$spc + par[3]*data$urban
  zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  zk <- par[7] + par[8]*data$midwest
  com <- par[9]*data$comp
  
  pik <- xb + zk
  piw <- xb + zw
  piduow <- xb + zw + com
  piduok <- xb + zk + com
  athrho = par[10]
  if (athrho>1){athrho=0.99}
  if (athrho<-1){athrho=-0.99}
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_duo - prob_w - prob_nofirm
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$y == 0, log(prob_nofirm), ifelse(data$y == 1, log(prob_w), ifelse(data$y == 2, log(prob_k), log(prob_duo))))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```

```{r}
optim(par=c(1.5, 1.5, 1.2, -11.8, -1.1, 0.72, -19.6, 0.37, -0.73, 0.32), walmartbi_ext_est, data_in=data, control = list("maxit"=10000))
```

# kmart moves first, estimate rho: 

we need to change the probability areas:

```{r}
kmart_est = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1]*data$population + par[2]*data$spc + par[3]*data$urban
  zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  zk <- par[7] + par[8]*data$midwest
  com <- par[9]*data$comp
  
  pik <- xb + zk
  piw <- xb + zw
  piduow <- xb + zw + com
  piduok <- xb + zk + com
  athrho = par[10]
  if (athrho>1){athrho=0.9}
  if (athrho<-1){athrho=-0.9}
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_k <- pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -pik, cov12 = athrho)
  prob_w <- 1 - prob_nofirm - prob_k - prob_duo
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$y == 0, log(prob_nofirm), ifelse(data$y == 1, log(prob_w), ifelse(data$y == 2, log(prob_k), log(prob_duo))))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```

```{r}
optim(par=c(1.5, 1.5, 1.2, -11.8, -1.1, 0.72, -19.6, 0.37, -0.7, 0.3), kmart_est, data_in=data, control = list("maxit"=10000))
```


# estimate firm specific delta 1 y delta 2

```{r}
delta_esp_est = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1]*data$population + par[2]*data$spc + par[3]*data$urban
  zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  zk <- par[7] + par[8]*data$midwest
  com1 <- par[10]*data$comp
  com2 <- par[11]*data$comp
  
  pik <- xb + zk
  piw <- xb + zw
  piduow <- xb + zw + com1
  piduok <- xb + zk + com2
  athrho = par[9]
  if (athrho>1){athrho=0.9}
  if (athrho<-1){athrho=-0.9}
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_nofirm - prob_w - prob_duo
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$y == 0, log(prob_nofirm), ifelse(data$y == 1, log(prob_w), ifelse(data$y == 2, log(prob_k), log(prob_duo))))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```

```{r}
optim(par=c(1.5, 1.5, 1.2, -11.8, -1.1, 0.72, -19.6, 0.37, -0.7, 0.3,-0.7,-0.7), delta_esp_est, data_in=data, control = list("maxit"=10000))
```


## 4 Nearly identical firms

rho=1 

```{r}
agg_noesp_rho1 = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1] + par[2]*data$population + par[3]*data$spc + par[4]*data$urban # this relates to entry (nfirms)
  #zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  #zk <- par[7] + par[8]*data$midwest
  com <- par[5]*data$comp # this to y...
  
  pik <- xb #+ zk
  piw <- xb #+ zw
  piduow <- xb + com #+ zw + com
  piduok <- xb + com #+ zk
  athrho = 0.99999 #par[6]
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_nofirm - prob_w - prob_duo
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$nfirms == 0, log(prob_nofirm), ifelse(data$nfirms == 1, log(prob_w) + log(prob_k), log(prob_duo)))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```


```{r}
optim(par=c(1,1.5, 1.5, 1.2, -.20), agg_noesp_rho1, data_in=data, control = list("maxit"=10000))
```


rho = 0.5


```{r}
agg_noesp_rho05 = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1] + par[2]*data$population + par[3]*data$spc + par[4]*data$urban # this relates to entry (nfirms)
  #zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  #zk <- par[7] + par[8]*data$midwest
  com <- par[5]*data$comp # this to y...
  
  pik <- xb #+ zk
  piw <- xb #+ zw
  piduow <- xb + com #+ zw + com
  piduok <- xb + com #+ zk
  athrho = 0.5 #par[6]
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_nofirm - prob_w - prob_duo
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$nfirms == 0, log(prob_nofirm), ifelse(data$nfirms == 1, log(prob_w) + log(prob_k), log(prob_duo)))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```


```{r}
optim(par=c(0, 1.5, 1.5, 1.2, -.20), agg_noesp_rho05, data_in=data, control = list("maxit"=10000))
```


### estimating rho nearly identical case

```{r}
agg_noesp_est_rho = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1] + par[2]*data$population + par[3]*data$spc + par[4]*data$urban # this relates to entry (nfirms)
  #zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  #zk <- par[7] + par[8]*data$midwest
  com <- par[5]*data$comp # this to y...
  
  pik <- xb #+ zk
  piw <- xb #+ zw
  piduow <- xb + com #+ zw + com
  piduok <- xb + com #+ zk
  athrho = par[6]
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_nofirm - prob_w - prob_duo
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$nfirms == 0, log(prob_nofirm), ifelse(data$nfirms == 1, log(prob_w) + log(prob_k), log(prob_duo)))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```


```{r}
optim(par=c(0, 1.5, 1.5, 1.2, -.20, 0.5), agg_noesp_est_rho, data_in=data, control = list("maxit"=10000))
```


# including Z

```{r}
agg_estZ = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1] + par[2]*data$population + par[3]*data$spc + par[4]*data$urban # this relates to entry (nfirms)
  zw <- par[5] +par[6]*data$dbenton + par[7]*data$southern
  zk <- par[8] + par[9]*data$midwest
  com <- par[10]*data$comp # this to y...
  
  pik <- xb #+ zk
  piw <- xb #+ zw
  piduow <- xb + com #+ zw + com
  piduok <- xb + com #+ zk
  athrho = par[11]
  if (athrho>1){athrho=0.9}
  if (athrho<-1){athrho=-0.9}
  
  prob_duo <- 1 - (pbinorm(-piduow, 100, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - (pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piduow, -piduok, cov12 = athrho)) - pbinorm(-piduow, -piduok, cov12 = athrho)
  
  prob_nofirm <- pbinorm(-piw, -pik, cov12 = athrho)
  prob_w <- pbinorm(100, -piduok, cov12 = athrho) - pbinorm(-piw, -piduok, cov12 = athrho)
  prob_k <- 1 - prob_nofirm - prob_w - prob_duo
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$nfirms == 0, log(prob_nofirm), ifelse(data$nfirms == 1, log(prob_w) + log(prob_k), log(prob_duo)))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```

```{r}
optim(par=c(1.5, 1.5, 1.2, -1.1, .72, -11.8, .37, -19.6, -0.2,-0.2, 0.5), agg_estZ, data_in=data, control = list("maxit"=10000))
```

## Univariate Case

```{r}
agg_noesp_univariate = function(par,data_in=data){
  # par -> c(a1,a2,a3,b0,b1,b2,c0,c1,d1)
  xb <- par[1] + par[2]*data$population + par[3]*data$spc + par[4]*data$urban # this relates to entry (nfirms)
  #zw <- par[4] +par[5]*data$dbenton + par[6]*data$southern
  #zk <- par[7] + par[8]*data$midwest
  com <- par[5]*data$comp # this to y...
  
  pik <- xb #+ zk
  piw <- xb #+ zw
  piduow <- xb + com #+ zw + com
  piduok <- xb + com #+ zk
  
  prob_duo <- 1 - (dnorm(-piduow) - dnorm(max(-piduow, -piduok))) - (dnorm(-piduok) - dnorm(max(-piduow, -piduok))) - dnorm(max(-piduow, -piduok))
  
  prob_nofirm <- dnorm(max(-piw, -pik))
  prob_w <- dnorm(-piduok) - dnorm(max(-piw, -piduok))
  prob_k <- 1 - prob_nofirm - prob_w - prob_duo
  prob_duo[prob_duo<0] <- 0
  prob_nofirm[prob_nofirm<0] <- 0
  prob_w[prob_w<0] <- 0
  prob_k[prob_k<0] <- 0
  
  lnf <- ifelse(data$nfirms == 0, log(prob_nofirm), ifelse(data$nfirms == 1, log(prob_w) + log(prob_k), log(prob_duo)))
  lnf[is.infinite(lnf)] <- 0
  return(-sum(lnf))
}
```


```{r}
optim(par=c(1.6417959,1.5124064, -0.9862969, 1.689094, 0.9191), agg_noesp_univariate, data_in=data, control = list("maxit"=10000))
```

